Prediction Assignment Writeup (Practical Machine Learning)
==========================================================

###Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

###Data 

The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 

(If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.) 

##Data Analysis

For this assignment we analyze the provided data to determine what activity an individual performed.

To do this, we should make use of caret and randomForest. This will allow us to generate correct answers for each of the 20 test data cases provided in the assignment. 

Here, we can make use of a seed value for consistent results.

```{r}
library(Hmisc)
library(caret)
library(randomForest)
library(foreach)
library(doParallel)
set.seed(2048)
options(warn=-1)
```

First, we load the data both from the provided training and test data provided by Coursera.

Some values contain a "#DIV/0!", which needs to be replaced with an "NA" value.

```{r}
training_data <- read.csv("pml-training.csv", na.strings=c("#DIV/0!") )
evaluation_data <- read.csv("pml-testing.csv", na.strings=c("#DIV/0!") )
```

We can also cast all columns from number 8 to the end to be numeric.

```{r}
for(i in c(8:ncol(training_data)-1)) {training_data[,i] = as.numeric(as.character(training_data[,i]))}

for(i in c(8:ncol(evaluation_data)-1)) {evaluation_data[,i] = as.numeric(as.character(evaluation_data[,i]))}
```

Some columns end up being primarily blank, which does not contribute well to the prediction.

Therefore, we can choose a feature set that only includes complete columns. We also should remove user name, timestamps and windows.

From here, we can determine and display out the feature set.

```{r}
feature_set <- colnames(training_data[colSums(is.na(training_data)) == 0])[-(1:7)]
model_data <- training_data[feature_set]
feature_set
```

We now have the model data built from our feature set.

```{r}
idx <- createDataPartition(y=model_data$classe, p=0.75, list=FALSE )
training <- model_data[idx,]
testing <- model_data[-idx,]
```

Next, we build 5 random forests with 150 trees each. We should make use of parallel processing to build this model. 

Utilizing several examples of how to perform parallel processing with random forests in R will help speed-up the process.

```{r}
registerDoParallel()
x <- training[-ncol(training)]
y <- training$classe

rf <- foreach(ntree=rep(150, 6), .combine=randomForest::combine, .packages='randomForest') %dopar% {
randomForest(x, y, ntree=ntree)
}
```

Finally, we should be sure to provide error reports for both training and test data.

```{r}
predictions1 <- predict(rf, newdata=training)
confusionMatrix(predictions1,training$classe)

predictions2 <- predict(rf, newdata=testing)
confusionMatrix(predictions2,testing$classe)
```

Conclusions and Submission of Test Data
---------------------------------------

As can be seen from the confusion matrix, this model is very accurate. There was no experimentation with PCA and other models. With test data around 99% accurate, it was expected that nearly all of the submitted test cases would be correct. In the end, they turned out to all be correct.

The following prepares the submission, by using the code provided in the instructions.

```{r}
pml_write_files = function(x){
 n = length(x)
 for(i in 1:n){
 filename = paste0("problem_id_",i,".txt")
 write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
 }
}


x <- evaluation_data
x <- x[feature_set[feature_set!='classe']]
answers <- predict(rf, newdata=x)

answers

pml_write_files(answers)
```